HO unification from object language to meta language

### 2024-04-02 paper/main.tex
notes on 8

TODO: dire che preserviamo l'invariante che tutte le variable sono fully-applied

\section{Handling of \maybebeta}\label{sec:beta}

@@

\begin{invariant}[\linkbeta \rhs]
  The \rhs of any \linkbeta has the shape $X~t_1\dots t_n$ such that $X$ is a
  flexible variable and $t_1\dots t_n$ is not in \llambda.
\end{invariant}

\linkbeta are put in \linkStore and activated when \rhs falls in \llambda.

% This property can be relaxed, to accept approximations: a dedicated
% section is given in a future section
\begin{corollary}
  If the \lhs of a \linkbeta is instantiated to a rigid term and its \rhs
  counterpart is still in \maybebeta, the current unification problem is not in
  \llambda and the unification fails.
\end{corollary}

@@

The list \elpiIn{Ag} is split into the list \elpiIn{Pf} and \elpiIn{Extra} such
that \elpiIn{append Pf Extra Ag} and \elpiIn{Pf} is the largest prefix of
\elpiIn{Ag} such that \elpiIn{Pf} is in \llambda. The \rhs of the \linkbeta is
the application of a fresh variable \elpiIn{C} having in scope all the free
variables appearing in the compiled version of \elpiIn{Pf} and \elpiIn{Extra}. The
variable \elpiIn{B}, returned has the compiled term, is a fresh variable having in
scope all the free variables occurring in \elpiIn{Pf1}.

\newcommand{\rhsBeta}{\ensuremath{N_S\appsep L}\xspace}

@@

TODO: Dire che maybe eta fa il detect anche su 
termini che non sono il llambda, oppure dirlo in section of maybeeta + dare un
esempio?

@@

\newcommand{\progBetaLL}{progress-beta-llambda}
\newcommand{\progBetaRH}{progress-beta-rigid-head}

@@

\begin{definition}[\progBetaLL]
  A link \linkbetaM{\Gamma}{X}{N_S\appsep [t|L]} is removed from \linkStore, if
  it exists a substitution $\gamma$, such that the first term $t$ is a name no
  occurring in $S$. In this case, let $M$ a fresh variable, then the new link
  \linkbetaM{\Gamma,x}{X}{M_{S,n}\appsep L} is added to \linkStore and if $L =
  []$, then $X$ is unified with $M_{S,n}$, otherwise, the
  \linketaM{\Gamma}{N}{\lambda x.M_x} is added to \linkStore.
  \label{def:progBetaLL}
\end{definition}

@@

\begin{proof}[Proof sketch]
  By \cref{def:progBetaRH}, the \linkbeta is removed from \linkStore, hence 
  they cannot be applied indefinitely. If the \linkbeta is progressed due to
  \progBetaLL, then two links are added, the first 
\end{proof}

\begin{corollary}
  Given a \linkbeta, the scope of its \rhs variables in \llambda.
\end{corollary}

>>>>

\todo[inline]{dire che preserviamo l'invariante che tutte le variable sono fully-applied
}
\section{Handling of \maybebeta}\label{sec:beta}

@@

\begin{invariant}[\linkbeta \rhs]
  The \rhs of any \linkbeta has the shape $X~t_1\dots t_n$ such that $X$ is a
  flexible variable and $t_1\dots t_n$ is not in \llambda.
  \todo[inline]{si dice: flexible term; oppure unification variable}
\end{invariant}

\linkbeta are put in \linkStore and activated when \rhs falls in \llambda. 
\todo[inline]{un po' ridondante con 8.2, non so se serva}

% This property can be relaxed, to accept approximations: a dedicated
% section is given in a future section
\begin{corollary}
  If the \lhs of a \linkbeta is instantiated to a rigid term and its \rhs
  counterpart is still in \maybebeta,
  the current unification problem is not in \llambda
  \todo[inline]{non capisco cosa vuoi dire con "the current...". Data la prova sotto vuoi dire che il problem originale sarebbe stato fuori... e quindi falliamo}
  and the unification fails.
\end{corollary}

@@

The list \elpiIn{Ag} is split into the list \elpiIn{Pf} and \elpiIn{Extra} such
that \elpiIn{append Pf Extra Ag} and \elpiIn{Pf} is the largest prefix of
\elpiIn{Ag} such that \elpiIn{Pf} is in \llambda. The \rhs of the \linkbeta is
the application of a fresh variable \elpiIn{C} having in scope all the free
variables appearing in the compiled version of \elpiIn{Pf} and \elpiIn{Extra}. The
variable \elpiIn{B}, returned has the compiled term, is a fresh variable having in
scope all the free variables occurring in \elpiIn{Pf1}
\todo[inline]{anche in Extra1 no?}.

\newcommand{\rhsBeta}{\ensuremath{N_S\appsep L}\xspace}
\todo[inline]{Io direi $X_{s_1\ldots s_n}\appsep l_1\ldots l_m$
... where S = [s1 .. sn] e L = [l1 .. lm]}

@@

\todo[inline]{Dire che maybe eta fa il detect anche su 
termini che non sono il llambda, oppure dirlo in section of maybeeta + dare un
esempio?}

@@

\newcommand{\progBetaLL}{\emph{progress-beta-\llambda}}
\newcommand{\progBetaRH}{\emph{progress-beta-rigid-head}}

@@

\begin{definition}[\progBetaLL]
  A link \linkbetaM{\Gamma}{X}{N_S\appsep [t|L]} is removed from \linkStore, if
  it exists a substitution $\gamma$, such that the first term $t$ is a name not
  occurring in $S$. In this case, let $M$ a fresh variable, then the new link
  \linkbetaM{\Gamma,x}{X}{M_{S,n}\appsep L} is added to \linkStore and if $L =
  []$, then $X$ is unified with $M_{S,n}$, otherwise, the
  \linketaM{\Gamma}{N}{\lambda x.M_x} is added to \linkStore.
  \label{def:progBetaLL}
  \todo[inline]{non è chiaro. la sostituzione immagino sia la corrente, non se ne esiste una. Poi non è chiaro se otherwise si riferische a in this case. Usa un enumerate}
\end{definition}

@@

\begin{proof}[Proof sketch]
  By \cref{def:progBetaRH}, the \linkbeta is removed from \linkStore, hence 
  they cannot be applied indefinitely. If the \linkbeta is progressed due to
  \progBetaLL, then two links are added, the first
  \todo[inline]{finire}
\end{proof}

\begin{corollary}
  Given a \linkbeta, the scope of its \rhs variables in \llambda.
  \todo[inline]{vuoi dire: le variabili che occorrono nel rhs sono in...?}
\end{corollary}
